---
title: "case study"
date: "2023-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load dataset:

```{r message=FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
data(nyc_airbnb)
```

```{r}
nyc_airbnb =
  nyc_airbnb |>
  rename(borough = neighbourhood_group) |>
  mutate(stars = review_scores_location / 2)
```

Questions we can ask:
- Where are Airbnbs most expensive?
    - Do borough, neighborhood, rating affect price?
- How long are Airbnb's active?
- Which units have the most availability?
- How does review impact impacted by location?
- How many apartments are run only by one host?
    - Does that affect price?
    - Does that vary by neighborhood?
- Where are the most Airbnbs?

Where are the most Airbnbs?
```{r}
nyc_airbnb |>
  count(borough)
```

Where are Airbnbs expensive? Which room types are most expensive?
(can replace mean with median if the distribution may be skewed)
```{r}
nyc_airbnb |>
  group_by(borough) |>
  summarize(mean_price = mean(price))

nyc_airbnb |>
  group_by(borough, room_type) |>
  summarize(mean_price = mean(price)) |>
  pivot_wider(
    names_from = room_type,
    values_from = mean_price
  )
```

Some visualizations:
```{r}
nyc_airbnb |>
  filter(price < 1000, room_type == "Entire home/apt") |>
  ggplot(aes(x = price)) +
  geom_histogram() +
  facet_grid(. ~ borough)
```

Where in Manhattan are places most expensive?
```{r}
nyc_airbnb |>
  filter(
    price < 1000,
    room_type == "Entire home/apt",
    borough == "Manhattan"
  ) |>
  group_by(neighbourhood) |>
  summarize(
    n_apts = n(),
    mean_price = mean(price)) |>
  arrange(desc(mean_price))
```

Visualizing distribution of price across Manhattan
- fct_reorder: take the average price of each neighborhood and re-order
```{r}
nyc_airbnb |>
  filter(
    price < 1000,
    room_type == "Entire home/apt",
    borough == "Manhattan"
  ) |>
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) |>
  ggplot(aes(x = price, y = neighbourhood)) +
  geom_density_ridges()
```

How many apartments are run only by one host?

How many units have the most and least availability?
```{r}
nyc_airbnb |>
  filter(availability_365 == 365 | availability_365 == 1) |>
  group_by(availability_365) |>
  count()
```

# Try to make a map

We have latitude and longitude

```{r}
nyc_airbnb |>
  filter(
    price < 1000,
    room_type == "Entire home/apt",
    borough == "Manhattan"
  ) |>
  ggplot(aes(x = lat, y = long, color = price)) +
  geom_point()
```

Let's try something else with the leaflet package:

- sample_n() lets you select a limited sample
- addProviderTiles(providers$CartoDB.Positron) to change the background to black and white
- addTiles() to be the standard colorful map
- need to install webshot or webshot2 package for knitr to take a screenshot
```{r}
library(leaflet)

nyc_airbnb |>
  filter(
    price < 1000,
    room_type == "Entire home/apt",
    borough == "Manhattan"
  ) |>
  sample_n(200) |>
  leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addMarkers(~lat, ~long)
```