---
title: "Strings and factors"
date: "2023-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(tidyverse)
library(p8105.datasets)
library(rvest)
```

# str_detect

str_detect tells you whether or not the item was found in each value of your string vector

str_replace(vector_name, "old", "new") to replace value

```{r}
string_vec = c("my", "name", "is", "sarah")

str_detect(string_vec, "sarah")
str_detect(string_vec, "e")

str_replace(string_vec, "sarah", "Sarah")
```

for exact matches, add symbols to the beginning or end of a line inside str_detect
  start with: ^ at the beginning
  ends with: $ at the end

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")
```

designate a list of characters that will count as a match with []

```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Pp]umpkin")
```

you don’t have to list these; instead, you can provide a range of letters or numbers that count as a match
  any instance that matches a number could be "[0-9]"
  add [a-z] so anything after is a lowercase letter
  
spaces are a character so you have to be explicit about allowing spaces

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-z]")
```

the character . matches everything

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

[], (), and . are special characters, so if you want to detect them, you have to use \\ before
  you can't use only 1 \ because \ is a special character
  if you want to detect [ use "\\["
  if you want to detect [# then use "\\[[0-9]"

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

this is why people don't recommend using special characters, e.g., spell out and instead of &

# Factors

vec_sex looks like a string variable so you need to specify it as a factor

use as.numeric to see the underlying numeric values

```{r}
vec_sex = factor(c("male", "male", "female", "female"))

as.numeric(vec_sex)
```

to specify male as first:

```{r}
vec_sex = fct_relevel(vec_sex, "male")
vec_sex
```

# NSDUH data

enter the url, read the html with the url, specify the table element, specify the first table, and skip the first row

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

marijuana_table =
  read_html(nsduh_url) |>
  html_table() |>
  first() |>
  slice(-1)
```

need to tidy this data set:
- the columns that contain P Value are nut helpful
- you have column names that are variables via pivot longer but do not pivot state (use -State)
- separate age & year by ( but put \\ before and then replace the ) with a space
- need to remove random a and b letters (that reference the footnotes)
- filter OUT certain regions using filter(!())

```{r}
marijuana_df =
  marijuana_table |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) |>
  separate(age_year, into = c("age", "year"), "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-b]", ""),
    percent = as.numeric(percent)
  ) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

plot & reorder state by lowest to highest percent using fct_reorder

```{r}
marijuana_df |>
  filter(age == "18-25") |>
  mutate(State = fct_reorder(State, percent)) |>
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# NYC Restaurant Inspections

```{r}
data("rest_inspec")
```

```{r}
rest_inspec |>
  count(boro, grade) |>
  pivot_wider(
    names_from = "grade",
    values_from = "n"
  ) |>
  knitr::kable()
```

```{r}
inspection_df =
  rest_inspec |>
  filter(grade %in% c("A", "B", "C"), boro != "Missing") |>
  mutate(
    boro = str_to_title(boro),
    dba = str_to_upper(dba)
  )
```

Pizza names

```{r}
inspection_df |>
  filter(str_detect(dba, "PIZZA")) |>
  count(boro)
```

reorder by frequency of boro using mutate(boro = fct_infreq(boro))
unlike color, fill shows by gradient

```{r}
inspection_df |>
  filter(str_detect(dba, "PIZZA")) |>
  mutate(boro = fct_infreq(boro)) |>
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```

order matters: replacing Manhattan's name after fct_infreq ignored fct_infreg

factors have very specific values, so trying to use a value that is not an existing factor level won’t work or using a standard non-forcats function like recode or rename would not work

```{r}
inspection_df |>
  filter(str_detect(dba, "PIZZA")) |>
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Manhattan", "The City")) |>
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```

instead, use fct_recode:

```{r}
inspection_df |>
  filter(str_detect(dba, "PIZZA")) |>
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")) |>
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```