---
title: "Data Wrangling II"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(rvest)
library(httr)
```

# Scraping

```{r message=FALSE}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
drug_use_html = read_html(url)

drug_use_html
```

now we will pull out relevant elements starting with the table using html_table()
- there are 15 tables stored as lists, but let's say we want the 1st table only using first()
- recommend removing rows by slicing (first row is a note)

```{r}
marijuana_df =
  drug_use_html |>
  html_table() |>
  first() |>
  slice(-1)
```

# CSS Selector

now we will construct a dataset for StarWars movies

```{r}
starwars_url = "https://www.imdb.com/list/ls070150896/"

starwars_html = read_html(starwars_url)
```

the data isn't stored in a table, so we will to figure out the elements we need using CSS Selector
- everything you want is highlighted with green or yellow, everything you don't want is green
- html_text() after html_elements() to extract the text

```{r}
starwars_movies =
  starwars_html |>
  html_elements(".lister-item-header a") |>
  html_text()

starwars_revenue =
  starwars_html |>
  html_elements(".text-small:nth-child(7) span:nth-child(5)") |>
  html_text()
```

# APIs

get water data from NYC's open data
- on the website, get data and click API Docs (convert to csv, not json)
- parse to clean it up

```{r}
nyc_water_df =
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") |>
  content("parsed")
```

new dataset:
- by default, the CDC API limits data to the first 1000 rows, so increase that by changing an element of the API query

```{r}
brffs_df =
  GET("https://data.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000)) |>
  content("parsed")
```

sometimes APIs don't give you structured things:

```{r}
poke_df = 
  GET("https://pokeapi.co/api/v2/pokemon/ditto") |>
  content()
```