---
title: "Tidy data"
date: "2023-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PULSE data

Create a pulse dataframe.

Haven package has functions for reading data in alternative formats.

```{r}
library(tidyverse)
pulse_df = 
  haven::read_sas("./public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi_score",
    names_prefix = "bdi_score_"
  ) |>
  mutate(
    visit = replace(visit, visit == "bl", "00m")
  )

view(pulse_df)

```

pivot_longer combined bl to 12m bdi scores into one column, "visit."

^names_prefix will get rid of that part of the prefix

```{r}
pulse_df_og = 
  haven::read_sas("./public_pulse_data.sas7bdat") |>
  janitor::clean_names()
```

Learning assessment:

Inside pivot longer, tell it the columns you need to pivot, the names it goes to, and values go to:

```{r}
litters_df = 
  read_csv("./FAS_litters.csv") |>
  janitor::clean_names() |>
  select(litter_number, gd0_weight, gd18_weight) |>
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd",
    values_to = "weight",
  ) |>
  mutate(
    gd = case_match(
      gd,
      "gd0_weight" ~ 0,
      "gd18_weight" ~ 18
    )
  )
```

# Bind rows

Import LoTR words data:

Specify columns by adding range argument to read_excel function
- Mutate added a column called movie
- Bring all of them together with bind
- Move from wide to long format since male and female are two columns but should be one and move their values to a new column called word
- Relocate moves that column to the front

```{r}
fellowship_df = 
  readxl::read_excel("./LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship")

two_towers_df = 
  readxl::read_excel("./LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two towers")

return_of_king_df = 
  readxl::read_excel("./LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return of the king")

lotr_df =
  bind_rows(fellowship_df, two_towers_df, return_of_king_df) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "gender",
    values_to = "word"
  ) |>
  relocate(movie)
```

There are 2 variables in one column, group: level of treatment (low, mod, con) and day of treatment (day 7 or 8)

```{r}
litters_df_2 = 
  read_csv("./FAS_litters.csv") |>
  janitor::clean_names() |>
  mutate(wt_gain = gd18_weight - gd0_weight) |>
  select(litter_number, group, wt_gain) |>
  separate(group, into = c("dose", "day_of_tx"), 3)

pups_df =
  read_csv("./FAS_pups.csv") |>
  janitor::clean_names() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    )
  )

fas_df =
  left_join(pups_df, litters_df_2, by = "litter_number")
```