---
title: "Visualization: sessions 8-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r barbie theme, message = FALSE}
# install.packages("remotes")
remotes::install_github("MatthewBJane/ThemePark")

library(ThemePark)
head(themepark_themes)
```

```{r}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

# Class 1: Basic scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

You can also take/start with the df and pipe the scatterplot, and it will create exactly the same plot:

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Piping may make filtering easier. You don't have to create a separate df just for New York, for example:

```{r}
nyc_weather =
  weather_df |>
  filter(name == "CentralParkNY") |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

nyc_weather + geom_point()
```

# Fancy plot

color can be added to the entire plot in ggplot or to the points in geom_point

geom_smooth adds a smooth curve

se = FALSE removes standard error bars from the curve

alpha blending/shading make the points more transparent (0.3 = 30% solid 70% opaque)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3) +
  geom_smooth(se = FALSE)
```

Plot with facets:

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  facet_grid(. ~ name)
```

Another fancy plot:

size = precipitation (higher precipitation = larger size)

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.3, size = 0.5) +
  geom_smooth() +
  facet_grid(. ~ name)
```

Assigning specific colors to specific things:

Need to assign the color in geom_point, not ggplot because aesthetic mappings takes variables from your dataset and maps them onto colors
```{r}
weather_df |>
  filter(name == "CentralPark_NY") |>
  ggplot(aes(x = date, y = tmax)) +
  geom_point(color = "blue")
```

Hex plot:

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

Line plot:
```{r}
weather_df |>
  filter(name == "Molokai_HI") |>
  ggplot(aes(x = date, y = tmax)) +
  geom_line()
```

Can combine line plot with points:
```{r}
weather_df |>
  filter(name == "Molokai_HI") |>
  ggplot(aes(x = date, y = tmax)) +
  geom_line(alpha = 0.3) +
  geom_point(size = 0.3)
```

# Univariate plotting

Easiest starting point for this is a basic histogram:

```{r}
ggplot(weather_df, aes(x = tmax)) +
  geom_histogram()
```

fill argument fills in colors for histogram

position = "dodge" places bars for each group side-by-side - bars should avoid each other rather than stack up on each other (however, this can get difficult --> easier to use density plots)

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge", binwidth = 2)
```

Density plot:

adjust argument is similar to binwidth in histograms

if you over-smooth or under-smooth, you may miss the message you're trying to make

density plots show more distribution than box plots

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3, adjust = 2)
```

Basic boxplots:

```{r}
ggplot(weather_df, aes(y = tmax)) +
  geom_boxplot()
```

Can compare by group, e.g., name:

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) +
  geom_boxplot()
```

Violin plot:

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) +
  geom_violin()
```

Ridge plot:

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) +
  geom_density_ridges(scale = 0.9)
```

# Saving and embedding plots

```{r}
ggp_weather =
  ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5)

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 5)
```

```{r, fig.width = 12, fig.height = 12}
ggp_weather
```

You can also set options globally in the beginning after loading packages with this code:

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

fig.asp = aspect ratio

# Class 2: Same plot from last time

labs(x or y) renames axis on the graph
  labs(color) renames the legend in this case since the colors are the legend
  labs(title) adds a header at the top
  labs(caption) adds a caption at the bottom
  
scale_x_continuous(breaks = ) tells you where you want the tick marks on the x axis
  (labels = ) helps you label the x axis ticks
  can do scale_y_continuous
    (position = "right") moves the y axis to the right side of the graph
    (trans = "sqrt") transforms the data
    (limits = c(0, 30)) zooms into/scales the graph to fit 0-30 instead of the default -15 to 40 range

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp",
    color = "Location",
    title = "Temperature plot",
    caption = "The data was retreieved from moaa"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0 C", "15 C")
  ) +
  scale_y_continuous(
    position = "right",
    limits = c(0, 30)
  )
```

# Changing colors

there are a lot of different scales you can adjust under scale_color

scale_color_hue works well for categorical variables
  e.g.,
    + scale_color_hue(h = c(100, 300))

viridis package is good for colors
  option = "magma" because magma is one of the color scales besides viridis, for example
  discrete = TRUE relates to variable type (categorical vs. continuous)

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp",
    color = "Location",
    title = "Temperature plot",
    caption = "The data was retreieved from moaa"
  ) +
  viridis::scale_color_viridis(discrete = TRUE)
```

other chart details:

theme_bw() inverts the default grey background to a white background
  be careful about the order in which you put this in - it can be a reset so try to put it high in order
  
  theme_classic() is like theme_bw() hides gridlines and a thick border
  
  theme_minimal() is my personal favorite

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp",
    color = "Location",
    title = "Temperature plot",
    caption = "The data was retreieved from moaa"
  ) +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()

nyc_weather_df =
  weather_df |>
  filter(name == "CentralPark_NY")

hawaii_weather_df =
  weather_df |>
  filter(name == "Molokai_HI")

ggplot(nyc_weather_df, aes(x = date, y = tmax)) +
  geom_point() +
  geom_line(data = hawaii_weather_df) +
  ThemePark::theme_barbie() +
  labs(
    title = "I made this barbie-themed plot for you"
  )
```

# Patchwork

faceting - I want this plot duplicated for another related variable
  not good if you want two completely different plots

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  facet_grid(. ~ name) +
  ThemePark::theme_barbie() +
  labs(
    title = "this one is slightly cuter"
  )
```

to put them side by side:

```{r}
ggp_temp_scatter =
  weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)

ggp_precip_density =
  weather_df |>
  ggplot(aes(x = prcp, color = name)) +
  geom_density()
```

# Data manipulation

factor variables are thought of as categorical variables with order/levels, whereas characters are just characters with no structure
  this is important now because when ggplot has to figure out what goes onto which axis and the order of colors, it converts it to a factor variable in the background and by default uses alphabetical order
  need to change the variable structure to change the other

fct_reorder function lets you put name in order of some other variable
  fct_relevel is manual?

```{r}
weather_df |>
  mutate(
    name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_MA"))
  ) |>
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

```{r}
weather_df |>
  mutate(
    name = fct_reorder(name, tmax)
    ) |>
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```

# Complex facet grid

```{r}
litters_df =
  read_csv("FAS_litters.csv") |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_treatment"), sep = 3)

pups_df =
  read_csv("FAS_pups.csv") |>
  janitor::clean_names()

FAS_df =
  left_join(pups_df, litters_df, by = "litter_number")

FAS_df |>
  select(dose, day_of_treatment, starts_with("pd")) |>
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "postnatal_day"
  ) |>
  drop_na() |>
  mutate(
    outcome =
      fct_reorder(outcome, postnatal_day)
  ) |>
  ggplot(aes(x = dose, y = postnatal_day)) +
  geom_violin() +
  facet_grid(day_of_treatment ~ outcome)
```

# Session 10

lubirdate: handy way of accessing month as a variable

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) |>
  select(name, id, everything())
```

```{r}
weather_df |>
  ggplot(aes(x = prcp)) +
  geom_histogram()
```

here are the big outliers:

```{r}
weather_df |>
  filter(prcp > 1000)
```

```{r}
weather_df |>
  filter(tmax >= 20, tmax <= 30) |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()
```

if you run this, you'll notice that grouping is sort of invisible
only hint below is # groups: name comment

(n_obs = n()) tells you to give you the number of observations
  good way to determine sample size in groups you care about

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(n_obs = n())
```

can also count number of names (and edit the name)

```{r}
weather_df |>
  count(name, name = "n_obs")
```


```{r}
weather_df |>
  count(name, month) |>
  pivot_wider(
    names_from = name,
    values_from = n
  )
```

# General summaries

can take us beyond simply counting

2 ways to remove NA:

```{r}
weather_df |>
  drop_na(tmax) |>
  group_by(name) |>
  summarize(
    mean_tmax = mean(tmax),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax)
  )

weather_df |>
  group_by(name) |>
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE),
    sd_tmax = sd(tmax, na.rm = TRUE)
  )
```

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |>
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

more pivot wider:

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |>
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  )
```

sometimes it's nice to format things as actual tables, especially if you're sending to others:
use knitr packages kable function

```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |>
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) |>
  knitr::kable(digits = 2)
```

# Grouped mutate

can compute the month-level mean instead of one mean for the entire dataset, for example

```{r}
weather_df |>
  group_by(name) |>
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax
  ) |>
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```

min_rank() gives the ranking from lowest to highest
  min_rank(desc()) puts it in descending/reverse order
filter for the coldest days in each month

```{r}
weather_df |>
  group_by(name, month) |>
  mutate(tmax_rank = min_rank(tmax)) |>
  filter(tmax_rank < 2)
```

lag(): for example, how does yesterday's temperature relate to today's temperature/how does previous temperature impact current/future temperatures
  lag(tmax, 3) would go 3 rows back
  if you forget your grouping, you will get random values

```{r}
weather_df |>
  group_by(name) |>
  mutate(
    yesterday_tmax = lag(tmax)
  )
```

show the day-to-day variation in temperature:

```{r}
weather_df |>
  group_by(name) |>
  mutate(temp_change = tmax - lag(tmax)) |>
  summarize(
    sd_temp_change = sd(temp_change, na.rm = TRUE)
  )
```